Topological Equivariant Artist Model for Visualization Library Architecture

### 2023-08-08 paper/main.tex
presheaf and coproduct diagrams

% *** diagrams
\usepackage{tikz}
\usetikzlibrary{cd, calc, positioning}
\usetikzlibrary{patterns,shapes.arrows}
\usepackage{pgfplots} %numberline
\usepgfplotslibrary{groupplots,dateplot}
\pgfplotsset{
  compat=1.15,
   simple axis/.style={
    scale only axis,
    axis x line=bottom,
    axis y line=none,
    axis line style={latex-latex},
    xticklabel style = {font=\footnotesize},
    width=40pt,
    height=20pt,
    xtick = {0, 180, 360},
    xticklabels = {$0$, $\pi$, $2\pi$},
    xmin = 0, xmax=360,
    ymin=0, ymax=1,
    domain=0:360,
    range=0:1,
    tickwidth=0,
    }
}
\pgfplotsset{compat=newest}

@@

\IEEEPARstart{V}{isualization} is a structure preserving mapping from data to graphic. One way of verifying that the structure of the data is present in the graphic is to verify that the data set can be distilled from the graph, i.e. that the values read off a bar chart are equivalent to the values used to construct the bar chart. Since structure preservation requires that the $data\leftrightarrow graphic$ be bidirectional, changes to either data or graphic are expected to propagate to the other side of the transformation\cite{kindlmannAlgebraicProcessVisualization2014}. In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics. We use category theory, which is the math of structures and their relations, to explicitly specify what it means for a visualization mapping to preserve structure in an explicit, comprehensive, and consistent way. We encode the structure of the data and graphics in a uniform generalizable manner by using a mathematical structure from algebraic topology called a fiber bundle; doing so allows us to discuss everything from lines to complex ND spaces with the same interface.

@@

For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
\begin{minipage}[c]{0.30\columnwidth} 
  \begin{equation}
    \label{eq:atct:base_union}
  \begin{tikzcd}
    \dbasec^{a}\sqcup_{\dbasec^{c}}\dbasec^{b} 
    & \dbasec_{b} \arrow[l, "\iota_b"', color=base] \\
    \dbasec_a \arrow[u, "\iota_a", color=base] 
    & \dbasec_c 
    \arrow[u, "\iota_c"', color=base] 
    \arrow[l, "\iota_c"', color=base] 
  \end{tikzcd}
\end{equation}
\end{minipage}\hfill
\begin{minipage}[c]{0.60\columnwidth}
  \begin{tikzpicture}[commutative diagrams/every diagram]
    \draw node[anchor=north west, minimum width=30pt, minimum height=10pt] (kabc) {};
\begin{axis}[simple axis, at={(kabc)}]
  %\addplot [thick, smooth] {sin(2*pi*x)};
\end{axis}

\draw node[anchor=north east, right=55pt of kabc, minimum width=30pt, minimum height=10pt] (kb) {};
%\begin{axis}[simple axis,at={(kb)}]
%  %\addplot [thick, smooth] {sin(2*pi*x)};
%\end{axis}
\draw[at=(kb)][draw=black](0,0) -- (360,0);
\draw node[anchor=south west, below=of kabc, minimum width=30pt, minimum height=10pt] (ka) {};
\begin{axis}[simple axis,at={(ka)}]
  %\addplot [thick, smooth] {sin(2*pi*x)};
\end{axis}

\draw node[anchor=south east, below=of kb, minimum width=30pt, minimum height=10pt] (kc) {};
\begin{axis}[simple axis, at={(kc)}]
  %\addplot [thick, smooth] {sin(2*pi*x)};
\end{axis}

\path[commutative diagrams/.cd, every arrow, every label]
  (kc) edge[] (kb)
       edge[] (ka)
  (ka) edge[] (kabc)
  (kb) edge[] (kabc);

  \end{tikzpicture}

\end{minipage}

@@

For example, lets define presheaves $\sheaf_1, \sheaf_2$. These are maps from intervals $\openset_1, \openset_2$ to a set of functions $\Gamma_1, \Gamma_2$ that are continuous over that interval: 
\begin{equation*}
\begin{tikzcd}
  \textcolor{set}{\Gamma_2 = \textcolor{set}{\left\{\begin{array}{c} 
      \textcolor{section}{constant}\\ 
      \textcolor{section}{sin}\\
      \textcolor{section}{cos}\\
    \end{array}\right\}}}               
  \arrow[rr, "\iota^*", color=set] &  & 
  \textcolor{set}{\Gamma_1 = \left\{\begin{array}{c} 
    \textcolor{section}{constant}\\ 
    \textcolor{section}{sin}\\
    \textcolor{section}{cos}\\
    \textcolor{section}{tan}\\
  \end{array}\right\}} \\
  \textcolor{base}{\openset_2 = \left(0,2\pi\right)} 
  \arrow[u, "\sheafc_{1}", maps to, color=sheaf, shift right=4.5]
  &  & 
  \textcolor{base}{\openset_1=\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)} 
  \arrow[ll, "\iota"', color=base] 
  \arrow[u, "\sheafc_{2}"', maps to, color=sheaf, shift right=4.5]
  \end{tikzcd}
\end{equation*}
The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,1\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$. The generalization of this constraint is that data structures that contain continuous functions must support interpolating them over arbitrarily small subspaces.

>>>>

% *** diagrams
\usepackage{tikz}
\usetikzlibrary{cd}

@@

\IEEEPARstart{V}{isualization} is a structure preserving mapping from data to graphic, which loosely means that the visualization retains how the data values are positioned relative to each other and the mathematical properties of those values. One way of verifying that the structure is preserved is to attempt to distill
the data set from the graphic, i.e. check that the values read off a bar chart are equivalent to the values used to construct the chart. Structure preservation assumes that the $data\leftrightarrow graphic$ mapping is bidirectional, meaning changes to either data or graphic are expected to be reflected in the other\cite{kindlmannAlgebraicProcessVisualization2014}.

 In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics using topology and category theory.

 We use category theory because it is the math of structures and their relations and therefore provides a robust framework for defining structure and expressing the conditions under which structure would be preserved.

 We encode the structure of the data and graphics in a uniform generalizable manner ; doing so allows us to discuss everything from lines to complex ND spaces with the same interface.

@@

For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
\begin{minipage}[c]{0.25\columnwidth} 
  \centering
  \begin{equation}
    \label{eq:atct:base_union}
  \begin{tikzcd}[row sep=large]
    \dbasec^{a}\sqcup_{\dbasec^{c}}\dbasec^{b} 
    & \dbasec_{b} \arrow[l, "\iota_b"', color=base] \\
    \dbasec_a \arrow[u, "\iota_a", color=base] 
    & \dbasec_c 
    \arrow[u, "\iota_c"', color=base] 
    \arrow[l, "\iota_c"', color=base] 
  \end{tikzcd}
\end{equation}
\end{minipage}

@@

For example, lets define presheaves $\sheaf_1, \sheaf_2$. These are maps from intervals $\openset_1, \openset_2$ to a set of functions $\Gamma_1, \Gamma_2$ that are continuous over that interval: 
\note{Human readable before inline not after}

\begin{figure}[H]
  \includegraphics*[width=1\columnwidth]{figures/tex/presheaf.pdf}
  \caption*{presheafs $\sheaf_1, sheaf_2$ mapping from intervals to the trignometric functions defined over those intervals }
\end{figure}

The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,1\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$. The generalization of this constraint is that data structures that contain continuous functions must support interpolating them over arbitrarily small subspaces.
